---
title: "coding6 phenomenological"
author: "Jen Richmond"
date: "02/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# load packages

```{r loadpackages, message=FALSE, warning=FALSE}

library(tidyverse)
library(here)
```

# read in the data

```{r readdata, message=FALSE, warning=FALSE}
phenom <- read_csv(here("data", "study3_phenom_ratings.csv"))
```

# filter df to only include participants with identifier = good

Some pp were excluded due to english skills, age, or not completing the experiment

```{r}

```


# familiarise yourself with the data

Use glimpse() or names() or str() to get an idea what is included in the dataset

```{r datastructure}

```

# details re phenomenological ratings

For each event that participants remembered/imagined they were asked to rate the following on a 7 point scale

- vividness
- pre/re-experiencing
  + the extent to which they felt like they were pre or re experiencing the event
- sensory details
  + individual sensory items (sight, sound, smell, taste) averaged into single sensory detail rating for each event
- contextual details
  + ratings of clarity of location and clarity of spatial arrangement of objects averaged into combined contextual details rating for each event
- perspective
  + ratio calculated by dividing rating for "seen through my own eyes" and "as if I was a fly on the wall" to determine field or observer perspective. Higher ratios indicate field perspective; lower ratios field perspective
- reaction
  + combined reaction score calculated by averaging intensity and physical reactions
- coherence
- importance
  
All phenomenological ratings averaged across event type (i.e. future neutral, future positive) for analysis. 

# make the data long

You want to end up with a dataset that has separate variables for event, valence, direction, rating type, rating
```{r}

```

# make a dataset that averages scores for each rating type across events

each participant came up with two events for each condition, rating scores were averaged across those two events. 

```{r}

```


# Plot away!!!

Questions to answer...

- Do pp with anxiety rate their experience of remembering past events or imagining future events differently to controls?
- Does it make a difference whether they were asked to remember/imagine positive, negative or neutral events?
- Do pp with anxiety rate future events as more important?
- Are pp with anxiety more likely to see future events from a fly on the wall perspective?

Other questions you may have...... :)
