---
title: "coding8 stats"
author: "Jen Richmond"
date: "02/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Framing questions to be answered with statistics

Now that you have visualised the specificity scores, details, and phenomenological scores, we need to decide what questions you want to answer with statistics. Refer to your plots for each DV and under the headings below write questions that you want to use statistical tests to answer (i.e. is there is a significant difference between ..., is there a correlation between....)

# specificity codes


# details


# phenomenological ratings







# load packages

```{r loadpackages, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(ggeasy)
library(ggbeeswarm)
library(RColorBrewer)
library(papaja)

```

# read in the data

```{r readdata, message=FALSE, warning=FALSE}

```

# familiarise yourself with the data

Use glimpse() or names() or str() to get an idea what is included in the dataset

```{r datastructure}

```

