---
title: "Coding exercise 2"
author: "Jen Richmond"
date: "10/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this coding exercise, you will work with data from a study published in Psychological Science. 

Beall, A. T., Hofer, M. K., & Shaller, M. (2016). Infections and elections: Did an Ebola outbreak influence the 2014 U.S. federal elections (and if so, how)? Psychological Science, 27, 595-605.

# load packages

```{r loadpackages, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(papaja)
library(naniar) # re missing data
library(RColorBrewer)
library(ggpubr)

```


# read in the data

```{r readdata, message=FALSE, warning=FALSE}
beall <- read_csv(here("data", "2_Beall.csv"))
```

# familiarise yourself with the data

Use glimpse() or names() or str() to get an idea what is included in the dataset

```{r datastructure}

```


# clean up the data

The variables are inconsistently named (upper, lower, .). Use clean_names() to make them consistent. 

```{r cleannames}

```


The first couple of rows are also filled with NA (no data). Use slice(3:n()) to slice the data from row 3 until the end (n()). (you will end up with 65 obs)

```{r slice}


```


There are lots of variables in this data set that we wont need. Make a smaller data set that only includes month, day, two_weeks, voter_intention_index, and ebola_search_volume_index. 

```{r selectdata}



```


# Is there a correlation between voter intention and ebola?

Use geom_point to make a scatter plot the relation between voter intention and ebola. Use theme_apa and add an informative title. 


```{r scatterplot}


```

Have a look at the `ggpubr` package and stat_cor() and see if you can get a correlation coefficient to display on the plot
```{r withcorrelation}

```

# Plot timeframes

### just two weeks prior

Restrict the data to only 2 weeks prior to the outbreak and plot again. 
```{r twoweeks}


```

### september

Restrict to just the month of september and plot date againest voter intention

```{r sept}
# whole month



```

Restrict to just the last week september and plot date againest voter intention

```{r weeksept}

```

### october

Restrict to just October and plot date against voter intention
```{r oct}
# whole month


```
Resrict to just the first week of October and plot date againest voter intention

```{r weekoct}
# first week 

```


# Missing data

There is so much missing data in the voter intention index.  Take your data and select just the voter intention index and pipe it into naniar::vis_miss() to plot how much data is missing.  63% missing!!

```{r vismiss}

```

Is there that much missing data for Sept?  

```{r sept_vismiss}

```

What about October?
```{r oct_vismiss}

```


Look up the original paper and see whether they were analysing voter intention with correlation, or whether the analysis involved more variables. 
