---
title: "Coding exercise 2"
author: "Jen Richmond"
date: "10/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this coding exercise, you will work with data from a study published in Psychological Science. 

Beall, A. T., Hofer, M. K., & Shaller, M. (2016). Infections and elections: Did an Ebola outbreak influence
the 2014 U.S. federal elections (and if so, how)? Psychological Science, 27, 595-605.

# load packages

```{r loadpackages, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(papaja)
library(naniar) # re missing data
library(RColorBrewer)

```


# read in the data

```{r readdata, message=FALSE, warning=FALSE}
beall <- read_csv(here("data", "2_Beall.csv"))
```

# familiarise yourself with the data
Use glimpse() or names() or str() to get an idea what is included in the dataset

```{r datastructure}
glimpse(beall)
```


# clean up the data

The variables are inconsistently named (upper, lower, .). Use clean_names() to make them consistent. 

```{r cleannames}
beall_clean <- beall %>%
  clean_names() 
```


The first couple of rows are also filled with NA (no data). Use slice(3:n()) to slice the data from row 3 until the end (n()). (you will end up with 65 obs)

```{r}

beall_sliced <- beall_clean %>%
  slice(3:n())
```


There are lots of variables in this data set that we wont need. Make a smaller data set that only includes month, day, two_weeks, voter_intention_index, and ebola_search_volume_index. 

```{r selectdata}

beall_select <- beall_sliced %>%
  select(month, date, two_weeks_prior_to_outbreak_only, voter_intention_index, ebola_search_volume_index)

```


# Q1 Is there a relation between voter intention and ebola?

Use geom_point to make a scatter plot the relation between voter intention and ebola. Use theme_apa and add an informative title. 

```{r scatterplot}
beall_select %>%
  ggplot(aes(x = voter_intention_index, y = ebola_search_volume_index)) +
  geom_point() +
  theme_apa()

```

```{r}
beall_select %>%
  filter(two_weeks_prior_to_outbreak_only == 1) %>%
  ggplot(aes(x = voter_intention_index, y = ebola_search_volume_index)) +
  geom_point() +
  theme_apa()

```

# Month of september
```{r}
# whole month

beall_select %>%
  filter(month == 9) %>%
  ggplot(aes(x = date, y = voter_intention_index)) +
  geom_point() +
  theme_apa()

# last week 
beall_select %>%
  filter(month == 9) %>%
  filter(date > 23) %>%
  ggplot(aes(x = date, y = voter_intention_index)) +
  geom_point() +
  theme_apa()
```

# Month of October
```{r}
# whole month

beall_select %>%
  filter(month == 10) %>%
  ggplot(aes(x = date, y = voter_intention_index)) +
  geom_point() +
  theme_apa()

# last week 
beall_select %>%
  filter(month == 10) %>%
  filter(date < 8) %>%
  ggplot(aes(x = date, y = voter_intention_index)) +
  geom_point() +
  theme_apa()
```


# Missing data

There is so much missing data in the voter intention index. Take your data and select just the voter intention index and pipe it into naniar::vis_mis() to plot how much data is missing.  

```{r}
beall_select %>%
  select(voter_intention_index) %>%
  vis_miss()
```

