---
title: "coding5_mtt_details"
author: "Jen Richmond"
date: "16/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages

```{r}

```

# read in the data

study_3_details_mtt.csv in the data folder

```{r}

```

# use table() to check how many participants we have in each group

```{r}

```

# drop the variables that are proportions that were calculated in excel

You will create them again below. 

```{r}

```


# make the data long

We have two group and pp in each group described 2 events in each of the following categories

- positive future
- neutral future
- negative future
- positive past
- neutral past
- negative past

There are variables in the dataset for the code (specific categorical general), number of internal details, external details. The data is in wide format. 

Make the data long so there are variables for
- event (event1, event2)
- valence (pos, neu, neg)
- direction (past/future)
- detail type (internal/external)

```{r}

```


# create a new variable that combines event 1 and 2 "codes" into a score out of 6

```{r}

```

# calculate the proportion of internal for each event and average across event 1 and 2

```{r}

```

NOTE, you might like to separate analysis of the coding scores and the details into separate data frames. 

# Plot away!!!

Questions to start with....

- do pp with anxiety have lower scores on the event codes? (i.e. report fewer events that meet criteria for specific?
- when you look at just the events that meet criteria for specific, do participants with anxiety report fewer internal details
- does the pattern of results look different if you use proportions (you'll need to use mutate to make a new variable) vs. raw internal/external scores. 
- are there interesting patterns across direction/valence ?

